---
title: "CHILDES uncertainty terms query"
output: html_document
---

## Set up

```{r}
rm(list=ls())
library(magrittr)
library(stringr)
library(knitr)
library(tidyverse)
library(RMySQL)
library(lazyeval)
```

## Connect to the CHILDES db 

Code adapted from the `dplyr` databases [vignette](https://cran.r-project.org/web/packages/dplyr/vignettes/databases.html)

Note that prior to running this code, you need to make sure that the mysql server is running. In the terminal, run: `mysql.server start`

```{r}
db <- src_mysql(host = "localhost", user='root', password='mbrllngs12', dbname='childes')
```

Get the relevant tables from the database

```{r}
corpora_meta_data <- tbl(db, "corpora")
kid_meta_data <- tbl(db, "children")
words <- tbl(db, "words")
```

## Extract all utterances with question words

```{r}
# start the clock
ptm <- proc.time()

all_qs <- words %>% 
  filter(gloss %in% c("maybe", "sometimes", "know", "unsure", "certain", "uncertain")) %>% 
  collect(n = Inf) 

# Stop the clock
proc.time() - ptm
```

Clean up the data 

```{r}
days_in_month <- 30.436875

all_qs %<>% 
  select(gloss,speaker, mor, participants, gender, age, child, corpus, 
         sentgloss, utt_number, date) %>% 
  filter(speaker %in% c("MOT", "CHI", "DAD", "FAT")) %>% 
  mutate(age_months = round(age / days_in_month, 0),
         child_clean = ifelse(child == "--", corpus, child),
         speaker_clean = ifelse(speaker == "CHI", "child", "parent"),
         parent_type = ifelse(speaker == "MOT", "mother", "father"))

all_qs %<>% mutate(utt_number = as.numeric(utt_number))
```

Get max utterance number for each recording, so we can normalize. 

```{r}
tmp <- all_qs %>% 
  group_by(child_clean, age_months) %>% 
  mutate(utt_number = as.numeric(utt_number)) %>% 
  summarise(length_transcript = max(utt_number))

all_qs %<>% left_join(., tmp, by = c("child_clean", "age_months"))
```

Get the surrounding utterances for a measure of discourse context

```{r}
context_df <- all_qs %>% select(gloss, utt_number, corpus, date)
```

```{r, eval = F}
## todo: get this context query function to work! 
data_f <- context_df[1,]

query_context_fun <- function(data_f, range_val = 5) {
  # get utterance range
  utt_number <- data_f[["utt_number"]]
  lower <- utt_num - range_val
  upper <- utt_num + range_val
  
  # get date
  this_date <- data_f[["date"]]
  # get corpus
  this_corpus <- data_f[["corpus"]]
  
  # todo: standard eval stuff so we can pass strings to dplyr functions
  
  # query the db
  surrounding_utts <- words %>% 
    filter(corpus == "Bloom", utt_number >= lower, utt_number <= upper) %>% 
    collect(n = 10) 
}
```

Save as csv

```{r, eval = F}
write_csv(x = all_qs, "../data/childes_uncertainty_terms.csv")
```
