---
title: "Childes Questions DB Queries"
author: "Kyle MacDonald"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

## Set up

Clear workspace and load relevant packages

```{r}
rm(list=ls())
library(magrittr)
library(stringr)
library(knitr)
library(tidyverse)
library(dbplyr)
library(RMySQL)
```

## Connect to the CHILDES database 

Code adapted from the `dbplyr` databases [vignette](https://cran.r-project.org/web/packages/dbplyr/vignettes/dbplyr.html).

Note that prior to running this code, you need to make sure that the local mysql server is running. In the terminal, run: `mysql.server start`

```{r}
db <- DBI::dbConnect(RMySQL::MySQL(), 
  host = "localhost",
  user = "root",
  password = rstudioapi::askForPassword("Database password"),
  dbname  = "childes_km"
)

dbListTables(db)
```

Get the relevant tables from the database

```{r}
#corpora_meta_data <- tbl(db, "corpora")
# kid_meta_data <- tbl(db, "children")
words <- tbl(db, "words")
```

## Extract all utterances with question words

Note the use of the *collect()* function to actually return the results of our sql query. If we don't include the collect function, then the dplyr code will never "touch" the database.

```{r}
# create SQL query 
all_qs_db <- words %>% filter(line_terminator == "?")  

# show the query translated to SQL
all_qs_db %>% show_query()
    
# collect the results so we can write to file (this will take some time)
all_qs <- all_qs_db %>% collect(n = Inf) 
```

Create session ID for each transcript.

```{r}
# get the number of transcripts for each child
# this way we know how many id numbers to expect
n_transcripts <- all_qs %>% 
  select(child, length_transcript) %>% 
  distinct() %>%
  group_by(child) %>% 
  mutate(n_sessions = n(),
         session_id = 1:n_sessions)

# join n_sessions info with rest of data
all_qs %<>% left_join(., n_transcripts, by = c("child", "length_transcript", "n_sessions", "session_id"))
```

## Clean up the data 

Only keep distinct questions

```{r}
all_qs %<>% 
  group_by(child, speaker, utt_number, session_id) %>% 
  distinct(sentgloss, .keep_all = T)
```

```{r}
days_in_month <- 30.436875

all_qs %<>% 
  select(gloss, speaker, mor, participants, gender, age, child, corpus, 
         sentgloss, utt_number, act, length_transcript, n_sessions, session_id) %>% 
  mutate(age_months = round(age / days_in_month, 0),
         child_clean = ifelse(child == "--", corpus, child),
         speaker_clean = ifelse(speaker == "CHI", "child", "adult"),
         parent_type = case_when(
           speaker == "MOT" ~ "mother",
           speaker == "FAT" ~ "father",
           TRUE ~ "other_adult"
         )
  )
```

Clean up child variable. Need to fix this in the Chat parser.

```{r}
all_qs %<>% 
  ungroup %>% 
  mutate(child_clean = case_when(
    str_detect(child, "ale") ~ "alex",
    str_detect(child, "eth") ~ "ethan",
    str_detect(child, "lil") ~ "lily",
    str_detect(child, "nai") ~ "naima",
    str_detect(child, "vio") ~ "violet", 
    str_detect(child, "wil") ~ "william",
    str_detect(child, "n") ~ "naomi",
    TRUE ~ child
    )
  ) %>% 
  select(-child)
```

```{r}
all_qs %<>% 
  mutate(question_word = case_when(
    str_detect(gloss, "why") ~ "why", 
    str_detect(gloss, "what") ~ "what",
    str_detect(gloss, "where") ~ "where",
    str_detect(gloss, "who") ~ "who",
    str_detect(gloss, "^how") ~ "how",
    str_detect(gloss, "when") ~ "when",
    str_detect(gloss, "which") ~ "which",
    str_detect(gloss, "whose") ~ "whose",
    str_detect(gloss, "whom") ~ "whom",
    str_detect(gloss, "can") ~ "can",
    TRUE ~ "other"
  ),
  utt_number = as.numeric(utt_number)
  )
```

Get the total number of questions asked by adults and child for each session, so we can normalize later.

```{r}
n_questions <- all_qs %>% select(child_clean, speaker_clean, session_id, sentgloss, utt_number) 
  
n_questions %<>% 
  group_by(child_clean, speaker_clean, session_id) %>% 
  summarise(n_questions_transcript = n())

n_questions
```

Join the total number of questions with the rest of the data

```{r}
all_qs %<>% left_join(., n_questions, by = c("session_id", "child_clean", "speaker_clean"))
```

Save as csv

```{r, eval = F}
write_csv(x = all_qs, "../../data/queries/childes_questions.csv")
```
