---
title: "CHILDES-Q Visualization and Exploration"
author: "Kyle MacDonald"
output: html_document
---

## Set up

```{r, include = F}
knitr::opts_chunk$set(echo=T, warning=FALSE, message=FALSE,
                      fig.width=12, fig.height=8)
```

```{r}
rm(list=ls())
library(ggrepel); library(magrittr)
library(stringr); library(knitr)
library(tidyverse); library(directlabels); 
library(tidytext); library(forcats)
theme_set(ggthemes::theme_few())
```

## Read data

```{r}
all_qs <- read_csv("../../data/queries/childes_questions.csv")
```

## Descriptives

```{r}
all_qs %>% distinct(corpus) %>% kable()
```

How many children in dataset? 

```{r}
all_qs %>% 
  distinct(child_clean) %>% 
  nrow() %>% 
  paste("n children =", .)
```

Plot distribution of question types for adults and children.

```{r}
all_qs %>% 
  ungroup() %>% 
  group_by(speaker_clean, question_word, length_transcript) %>% 
  summarise(count = n()) %>%
  ggplot(., aes(x = question_word, y = count)) +
  geom_bar(stat = "identity") +
  labs(x = "Question Word") +
  facet_wrap(~speaker_clean)
```

How many questions do we have for adults and children?

```{r}
all_qs %>% 
  group_by(speaker_clean) %>% 
  distinct(sentgloss) %>% 
  summarise(count = n())
```

What are some examples of these questions? 

```{r}
all_qs %>% 
  filter(speaker_clean == "child") %>% 
  select(question_word, sentgloss, age_months) %>% 
  group_by(question_word) %>% 
  sample_n(., size = 1) %>% 
  knitr::kable()
```

## Exploratory visualizations

Histograms of question types for each child

```{r}
all_qs %>% 
  ungroup() %>% 
  filter(speaker_clean == "child") %>% 
  group_by(child_clean, question_word) %>% 
  summarise(freq = n()) %>% 
  ggplot(., aes(x = question_word, y = freq)) +
  geom_bar(stat = "identity") +
  labs(x = "Question Word", y = "Count",
       title = "Distribution of children's question types") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~child_clean, scales = "free_y")
```

## Word frequencies in question utterances

First, let's test the unnest_tokens() function on a subset of the data

```{r}
test_df <- all_qs %>% filter(child_clean == "gia")

test_output <- test_df %>% 
  unnest_tokens(word, sentgloss) %>% 
  left_join(test_df, .) %>% 
  select(sentgloss, word, utt_number)
```

Visual inspect of `test_output` suggests that the unnest_tokens() function is working properly, so we can apply it to the rest of the questions data. 

```{r}
unigrams_df <- all_qs %>% 
  unnest_tokens(word, sentgloss) %>% 
  select(-gloss)
```

Remove "0" from word column and filter blank rows. We are also going to remove "xxx" and "yyy" word codes.

```{r}
unigrams_df %<>% 
  mutate(word = str_replace_all(word, pattern = "0", "")) %>% 
  filter(word != "", word != "xxx", word != "yyy")
```

What are the most common words that occur in utterances flagged as questions?

```{r}
plot_qs <- unigrams_df %>% 
  group_by(word) %>% 
  count() %>%
  ungroup() %>% 
  top_n(30) %>% 
  arrange(desc(n)) %>%
  mutate(word = factor(word, levels = rev(unique(word))))

plot_qs %>% 
  mutate(word = reorder(word, n)) %>% 
  ggplot(aes(word, n)) +
  geom_col() +
  labs(x = NULL, y = "count") +
  coord_flip()
```

This isn't super informative. So now let's make the same plot but split by the canonical question words categories

```{r}
unigrams_df %>%
  group_by(question_word, word) %>% 
  count() %>% 
  top_n(5) %>% 
  mutate(word = reorder(word, n)) %>% 
  ggplot(aes(word, n, fill = question_word)) +
  geom_col(show.legend = FALSE) +
  labs(x = NULL, y = "count") +
  facet_wrap(~question_word, scales = "free") +
  coord_flip()
```

Remove stop words and do the same plot

```{r}
unigrams_df_filt <- unigrams_df %>% anti_join(stop_words) 

unigrams_df_filt %>%
  group_by(question_word, word) %>% 
  count() %>% 
  top_n(5) %>% 
  mutate(word = reorder(word, n)) %>% 
  ggplot(aes(word, n, fill = question_word)) +
  geom_col(show.legend = FALSE) +
  labs(x = NULL, y = "count") +
  facet_wrap(~question_word, scales = "free") +
  coord_flip()
```

Next, let's look at the most frequent words that occur together

## Bigram analyses

First, let's test the unnest_tokens() function is creating sensible bigrams

```{r}
test_df %>% 
  unnest_tokens(bigram, sentgloss, token = "ngrams", n = 2) %>% 
  left_join(test_df, .) %>% 
  select(sentgloss, bigram) %>% 
  head() %>% 
  kable()
```

Looks reasonable to me. let's do if for all the questions

```{r}
bigrams_df <- all_qs %>% 
  unnest_tokens(bigram, sentgloss, token = "ngrams", n = 2) %>% 
  filter(str_detect(bigram, "yyy") == F, 
         str_detect(bigram, "xxx") == F,
         str_detect(bigram, "0") == F) %>% 
  select(-gloss)
```

What are the most common bigrams that appear in questions for adults and kids?

```{r}
bigram_plot <- bigrams_df %>%
  group_by(bigram, speaker_clean) %>% 
  count() %>%
  group_by(speaker_clean) %>% 
  top_n(10) %>% 
  arrange(desc(n)) %>%
  ungroup %>% 
  mutate(bigram = factor(bigram, levels = rev(unique(bigram))))

bigram_plot %>% 
  ungroup %>%
  ggplot(aes(bigram, n, fill = speaker_clean)) +
  geom_col(show.legend = FALSE) +
  labs(x = NULL, y = "count") +
  coord_flip() +
  facet_wrap(~speaker_clean, scales = "free")
```

Plot relationship between adult and kid bigrams, removing transcripts with very few questions

```{r}
bigrams_cor_plot <- bigrams_df %>%
  filter(n_questions_transcript > 5) %>% 
  count(bigram, speaker_clean, child_clean, session_id) %>% 
  group_by(speaker_clean, child_clean, session_id) %>% 
  mutate(prop = n / sum(n))

ss <- bigrams_cor_plot %>% 
  group_by(bigram, speaker_clean, child_clean) %>% 
  summarise(m_prop = mean(prop, na.rm = T)) %>% 
  group_by(speaker_clean, bigram) %>% 
  summarise(m_prop = mean(m_prop)) %>% 
  spread(key = speaker_clean, value = m_prop)
```

```{r}
library(scales)

ss %>% 
  filter(complete.cases(.)) %>% 
  ggplot(aes(x = child, y = adult)) +
  geom_jitter(alpha = 0.1, size = 2.5, width = 0.3, height = 0.3) +
  geom_text(aes(label = bigram), check_overlap = TRUE) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  geom_abline(color = "gray40", lty = 2) 
```

## Most frequent bigrams for each child

```{r}
bigrams_child <- bigrams_df %>% 
  filter(n_questions_transcript > 5) %>%
  filter(speaker_clean == "child") %>% 
  count(bigram, child_clean) %>% 
  group_by(child_clean) %>%
  mutate(proportion = n / sum(n))
```

```{r}
bigrams_child %>% 
  filter(child_clean != "gia") %>% # gia doesn't have enoug data?
  group_by(child_clean) %>% 
  top_n(5) %>% 
  ggplot(aes(bigram, proportion, fill = child_clean)) +
  geom_col(show.legend = FALSE) +
  labs(x = NULL, y = "proportion of bigrams") +
  coord_flip() +
  facet_wrap(~child_clean, scales = "free")
```

Now do the same thing, but for each question word type

```{r}
bigrams_df %>% 
  filter(speaker_clean == "child") %>% 
  group_by(question_word) %>% 
  count(bigram) %>% 
  mutate(proportion = n / sum(n)) %>% 
  top_n(5) %>% 
  arrange(desc(proportion)) %>%
  mutate(bigram = factor(bigram, levels = rev(unique(bigram)))) %>% 
  ggplot(., aes(x = bigram, y = proportion, fill = question_word)) +
  geom_col(show.legend = F) +
  labs(x = NULL, y = "proportion of bigrams") +
  coord_flip() +
  facet_wrap(~question_word, scales = "free")
```

### Trigrams

```{r}
trigrams_df <- all_qs %>% 
  unnest_tokens(trigram, sentgloss, token = "ngrams", n = 3) %>% 
  filter(str_detect(trigram, "yyy") == F, 
         str_detect(trigram, "xxx") == F,
         str_detect(trigram, "0") == F) %>% 
  select(-gloss)
```

```{r}
trigram_plot <- trigrams_df %>%
  group_by(speaker_clean) %>%
  count(trigram) %>% 
  top_n(15, n) 
```

Now plot.

```{r}
trigram_plot %<>% 
  ungroup() %>%
  arrange(speaker_clean, n) %>% 
  mutate(order = row_number()) 
  
ggplot(trigram_plot, aes(order, n, fill = speaker_clean)) +
  geom_col(show.legend = FALSE) +
  labs(x = NULL, y = "count") +
  facet_wrap(~speaker_clean, scales = "free") +
  scale_x_continuous(
    breaks = trigram_plot$order,
    labels = trigram_plot$trigram,
    expand = c(0,0)
  ) +
  coord_flip() 
```

What about splitting the most frequent trigrams by question type? 

```{r}
trigrams_by_q <- trigrams_df %>%
  filter(speaker_clean == "child") %>% 
  group_by(question_word) %>% 
  count(trigram) %>% 
  top_n(5) %>% 
  ungroup

tri_plot_byq <- trigrams_by_q %>% 
  arrange(question_word, n) %>% 
  mutate(order = row_number()) 

ggplot(tri_plot_byq, aes(order, n, fill = question_word)) +
  geom_col(show.legend = FALSE) +
  labs(x = NULL, y = "count",
       title = "Most common trigrams in children's questions") +
  facet_wrap(~question_word, scales = "free") +
  scale_x_continuous(
    breaks = tri_plot_byq$order,
    labels = tri_plot_byq$trigram,
    expand = c(0.1,0)
  ) +
  coord_flip() +
  theme_bw()
```

Now do the same thing for adults' questions

```{r}
trigrams_by_q <- trigrams_df %>%
  filter(speaker_clean == "adult") %>% 
  group_by(question_word) %>% 
  count(trigram) %>% 
  top_n(5) %>% 
  ungroup

tri_plot_byq <- trigrams_by_q %>% 
  arrange(question_word, n) %>% 
  mutate(order = row_number()) 

ggplot(tri_plot_byq, aes(order, n, fill = question_word)) +
  geom_col(show.legend = FALSE) +
  labs(x = NULL, y = "count",
       title = "Most common trigrams in adults' questions") +
  facet_wrap(~question_word, scales = "free") +
  scale_x_continuous(
    breaks = tri_plot_byq$order,
    labels = tri_plot_byq$trigram,
    expand = c(0.1,0)
  ) +
  coord_flip() +
  theme_bw()
```

## Developmental analyses

Plot the proportion of utterances that were questions as a function of age.

```{r}
prop_qs_df <- all_qs %>% 
  filter(n_questions_transcript > 5) %>% 
  mutate(prop = n_questions_transcript / length_transcript) %>% 
  group_by(child_clean, age_months, speaker_clean) %>% 
  summarise(m = mean(prop, na.rm = T)) %>% 
  group_by(age_months, speaker_clean) %>% 
  summarise(m = mean(m, na.rm = T),
            n = n()) %>% 
  filter(is.na(age_months) == F)
```

Now plot

```{r}
prop_qs_df %>% 
  ggplot(., aes(x = age_months, y = m, size = n)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "loess", se = F, size = 2) +
  labs(x = "age (months)", y = "Prop. of Utterances",
       title = "Proportion of utterances that are questions") +
  langcog::scale_color_solarized() +
  theme(legend.position = "bottom") +
  facet_wrap(~speaker_clean, scales = "free_y")
```

Now we do the same analysis, but split by question word.

```{r}
dev_plot <- all_qs %>% 
  filter(n_questions_transcript > 5) %>% 
  count(question_word, child_clean, speaker_clean, session_id, n_questions_transcript,
        age_months) %>% 
  mutate(prop = n / n_questions_transcript) %>% 
  group_by(child_clean, age_months, question_word, speaker_clean) %>% 
  summarise(ss_prop = mean(prop, na.rm = T)) %>% 
  group_by(age_months, question_word, speaker_clean) %>% 
  summarise(m_prop = mean(ss_prop),
            n = n()) 
```

```{r}
dev_plot %>% 
  filter(speaker_clean == "child") %>% 
  ggplot(., aes(x = age_months, y = m_prop, color = question_word)) +
  geom_point(aes(size = n), alpha = 0.2) +
  geom_smooth(se=F, method = "loess", size = 2) +
  xlim(10, 68) +
  labs(y = "Prop. of Questions", x= "Age (months)",
       title = "Development of children's question asking") +
  guides(color = F) +
  langcog::scale_color_solarized() +
  facet_wrap(~question_word, scales = "free") +
  theme(legend.position = "top") +
  theme_bw()
```

### Analyze single word questions

```{r}
all_qs %<>% 
  rowwise() %>% 
  mutate(sentence_len = length(strsplit(sentgloss,' ')[[1]]))
```


```{r}
single_words <- all_qs %>% 
  filter(sentence_len == 1, n_questions_transcript > 5) %>% 
  count(question_word, child_clean, speaker_clean, session_id, n_questions_transcript,
        age_months) %>% 
  mutate(prop = n / n_questions_transcript) %>% 
  group_by(child_clean, age_months, question_word, speaker_clean) %>% 
  summarise(ss_prop = mean(prop, na.rm = T)) %>% 
  group_by(age_months, question_word, speaker_clean) %>% 
  summarise(m_prop = mean(ss_prop),
            n = n()) 
```

Plot

```{r}
single_words %>% 
  filter(speaker_clean == "child") %>% 
  ggplot(., aes(x = age_months, y = m_prop, color = question_word)) +
  geom_point(aes(size = n), alpha = 0.2) +
  geom_smooth(se=F, method = "loess", size = 2) +
  xlim(10, 68) +
  labs(y = "Prop. of Questions", x= "Age (months)",
       title = "Development of children's single word questions") +
  guides(color = F) +
  langcog::scale_color_solarized() +
  facet_wrap(~question_word, scales = "free") +
  theme(legend.position = "top") +
  theme_bw()
```

What are the single word questions in the other category?

```{r}
single_other <- all_qs %>% 
  filter(question_word == "other", sentence_len == 1)
```

```{r}
single_other_plot <- single_other %>% 
  count(gloss, speaker_clean) %>% 
  group_by(speaker_clean) %>% 
  top_n(15, n) %>% 
  ungroup() %>%
  arrange(speaker_clean, n) %>% 
  mutate(order = row_number()) 

ggplot(single_other_plot, aes(order, n, fill = speaker_clean)) +
  geom_col(show.legend = FALSE) +
  labs(x = NULL, y = "count") +
  facet_wrap(~speaker_clean, scales = "free") +
  scale_x_continuous(
    breaks = single_other_plot$order,
    labels = single_other_plot$gloss,
    expand = c(0,0)
  ) +
  coord_flip() 
```
