---
title: "Childes-q"
author: "Kyle MacDonald"
date: "2/20/2017"
output: html_document
---

## Set up

```{r}
rm(list=ls())
library(magrittr)
library(stringr)
library(knitr)
library(tidyverse)
library(RMySQL)
library(directlabels)
theme_set(ggthemes::theme_few())
```

## Connect to the CHILDES db 

Code adapted from the `dplyr` databases [vignette](https://cran.r-project.org/web/packages/dplyr/vignettes/databases.html)

Note that prior to running this code, you need to make sure that the mysql server is running. In the terminal, run: `mysql.server start`

```{r}
db <- src_mysql(host = "localhost", user='root', password='mbrllngs12', dbname='childes')
```

Get the relevant tables from the database

```{r}
corpora_meta_data <- tbl(db, "corpora")
kid_meta_data <- tbl(db, "children")
words <- tbl(db, "words")
```

## Extract all utterances with question words

```{r}
# start the clock
ptm <- proc.time()

all_qs <- words %>% 
  filter(gloss %in% c("why", "what", "where", "when", "which", "how", "who", "whose", "whom" )) %>% 
  collect(n = Inf) 

# Stop the clock
proc.time() - ptm
```

Clean up the data 

```{r}
days_in_month <- 30.436875

all_questions <- all_qs %>% 
  select(gloss,speaker, mor, participants, gender, age, child, corpus, 
         sentgloss, utt_number) %>% 
  filter(speaker %in% c("MOT", "CHI", "DAD", "FAT")) %>% 
  mutate(age_months = round(age / days_in_month, 0),
         child_clean = ifelse(child == "--", corpus, child),
         speaker_clean = ifelse(speaker == "CHI", "child", "parent"),
         parent_type = ifelse(speaker == "MOT", "mother", "father"))
```

```{r}
all_questions %<>% 
  mutate(gloss_clean = ifelse(str_detect(gloss, "why"), "why", 
                              ifelse(str_detect(gloss, "what"), "what",
                                     ifelse(str_detect(gloss, "where"), "where",
                                            ifelse(str_detect(gloss, "who"), "who",
                                                   ifelse(str_detect(gloss, "how"), "how",
                                                          ifelse(str_detect(gloss, "when"), "when",
                                                                 ifelse(str_detect(gloss, "which"), "which",
                                                                        ifelse(str_detect(gloss, "whose"), "whose",
                                                                               ifelse(str_detect(gloss, "whom"), "whom",
                                                                               "other"))))))))),
         utt_number = as.numeric(utt_number))
```


Get max utterance number for each recording, so we can normalize. 

```{r}
tmp <- all_questions %>% 
  group_by(child_clean, age_months) %>% 
  mutate(utt_number = as.numeric(utt_number)) %>% 
  summarise(length_transcript = max(utt_number))

all_questions %<>% left_join(., tmp, by = c("child_clean", "age_months"))
```

Save as csv

```{r, eval = F}
write_csv(x = all_questions, "../data/childes_questions.csv")
```

## Descriptives

```{r}
all_questions %>% distinct(corpus) %>% kable()
```

```{r}
all_questions %>% 
  distinct(child) %>% 
  nrow() %>% 
  paste("n children =", .)
```

```{r}
all_questions %>% 
  ungroup() %>% 
  group_by(speaker_clean, gloss_clean, length_transcript) %>% 
  summarise(count = n()) %>% 
  ggplot(., aes(x = gloss_clean, y = count)) +
  geom_bar(stat = "identity") +
  labs(x = "Question Word") +
  facet_wrap(~speaker_clean)
```

```{r}
all_questions %>% 
  group_by(speaker_clean, parent_type) %>% 
  summarise(count = n())
```

```{r}
all_questions %>% 
  filter(speaker_clean == "child") %>% 
  select(gloss_clean, sentgloss, age_months) %>% 
  group_by(gloss_clean) %>% 
  sample_n(., size = 2) %>% 
  knitr::kable()
```

## Exploratory analyses

Count different kinds of questions for each kid and plot as a function of age.

```{r}
all_questions %>% 
  filter(speaker == "CHI") %>% 
  group_by(child_clean, age_months, gloss_clean, length_transcript) %>% 
  summarise(freq = n()) %>% 
  filter(freq > 0) %>% 
  mutate(prop = freq / length_transcript) %>%
  ggplot(., aes(x = age_months, y = prop, color = gloss_clean)) +
  geom_smooth(method = "loess", se = F) +
  labs(color = "question word", y = "Prop. of Utterances") +
  langcog::scale_color_solarized() +
  facet_wrap(~child_clean, ncol = 7) +
  theme(legend.position = "bottom") +
  ggtitle("Children's questions")
```

Count different kinds of questions for each parent and summarise over development

```{r}
all_questions %>% 
  filter(speaker_clean == "parent") %>% 
  group_by(child_clean, age_months, gloss_clean, length_transcript) %>% 
  summarise(freq = n()) %>% 
  filter(freq > 0) %>% 
  mutate(prop = freq / length_transcript) %>% 
  ggplot(., aes(x = age_months, y = prop, color = gloss_clean)) +
  geom_smooth(method = "loess", se = F) +
  labs(color = "question word") +
  langcog::scale_color_solarized() +
  facet_wrap(~child_clean, ncol = 7) +
  theme(legend.position = "bottom") +
  ggtitle("Parents' questions")
```

Now get the average frequency for each question type at each month of development.

```{r}
all_questions %>% 
  filter(gloss_clean %in% c('what', 'where', 'why')) %>% 
  group_by(child_clean, age_months, gloss_clean, speaker_clean, length_transcript) %>% 
  summarise(freq = n()) %>% 
  filter(freq > 0) %>% 
  mutate(prop = freq / length_transcript) %>% 
  group_by(age_months, gloss_clean, speaker_clean) %>% 
  summarise(m = mean(prop, na.rm = T),
            n = n()) %>% 
  ggplot(., aes(x = age_months, y = m, color = gloss_clean)) +
  geom_point(aes(size = n), alpha = 0.3) +
  geom_smooth(se=F, method = "loess", size = 2) +
  xlim(10, 68) +
  labs(y = "Prop. of Utterances", x= "Age (months)") +
  guides(color = F) +
  langcog::scale_color_solarized() +
  geom_dl(aes(label = gloss_clean), method=list("last.points")) +
  facet_wrap(~speaker_clean, scales = "free") +
  theme(legend.position = "top")
```

### Gender analysis

```{r}
gender_ms <- all_questions %>% 
  group_by(child_clean, gloss_clean, speaker, gender) %>% 
  summarise(freq = n()) %>%
  filter(freq > 0, gender %in% c("male", "female")) %>% 
  group_by(gloss_clean, speaker, gender) %>% 
  langcog::multi_boot_standard(column = "freq", empirical_function = "mean") 
```

```{r}
gender_ms %>% 
  ggplot(., aes(x = gloss_clean, y = mean, color = gender)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), 
                  position = position_dodge(width = 0.3), size = 0.8) +
  labs(y = "Avg. Freq", x= "Question Type") +
  langcog::scale_color_solarized() +
  facet_wrap(~speaker, scales = "free") +
  theme(legend.position = "top")
```

No gender differences in question words produced by kid or by mom.

### Context analyses

Here we will attempt to operationalize cost and reward of a behavior to get the value for a question.

Spread participants variable, so we can code for the number of people in the conversation.

```{r}

```

### Complexity of utterance that each question word occurs in

Use some string functions to get the word count for each utterance.

```{r}
all_questions %<>% 
  rowwise() %>% 
  mutate(sentence_len = length(strsplit(sentgloss,' ')[[1]]))
```

Now summarise the mean sentence length for each question type for each kid and speaker

```{r}
sentence_len_ms <- all_questions %>% 
  ungroup() %>% 
  group_by(age_months, speaker_clean, child_clean, gloss_clean) %>% 
  summarise(m.ss = median(sentence_len), 
            n = n()) %>% 
  group_by(age_months, speaker_clean, gloss_clean) %>% 
  summarise(m = mean(m.ss), 
            n = n()) 
```

```{r}
sentence_len_ms %>% 
  filter(gloss_clean %in% c("where", "what", "why")) %>% 
  ggplot(data = ., aes(x = age_months, y = m, color = gloss_clean)) +
  geom_point(aes(size = n), alpha = 0.3) +
  geom_smooth(se=F, method = "loess", size = 2) +
  xlim(10, 68) +
  labs(y = "Avg. Sentence Length", x = "Age (months)") +
  guides(color = F) +
  langcog::scale_color_solarized() +
  geom_dl(aes(label = gloss_clean), method=list("last.points")) +
  facet_wrap(~speaker_clean) +
  theme(legend.position = "top")
```

### Single question word utterances

```{r}
single_words <- all_questions %>% 
  filter(sentence_len == 1) %>% 
  group_by(speaker_clean, child_clean, gloss_clean, age_months, length_transcript) %>% 
  summarise(freq = n()) %>% 
  mutate(prop = freq / length_transcript) %>% 
  group_by(speaker_clean, gloss_clean, age_months) %>% 
  summarise(m.ss = median(prop), 
            n = n())
```

```{r}
single_words %>% 
  filter(gloss_clean %in% c("where", "what", "why")) %>% 
  ggplot(data = ., aes(x = age_months, y = m.ss, color = gloss_clean)) +
  geom_point(aes(size = n), alpha = 0.3) +
  geom_smooth(se=F, method = "loess", size = 2) +
  xlim(10, 68) +
  guides(size = F) +
  labs(y = "Prop. of Utterances", x = "Age (months)") +
  langcog::scale_color_solarized() +
  facet_wrap(~speaker_clean) +
  theme(legend.position = "top")
```

### Fathers compared to mothers on wh-questions

```{r}

```


### Other analyeses to do:

- 
