---
title: "CHILDES-Q Visualization"
author: "Kyle MacDonald"
date: "3/20/2017"
output: html_document
---


## Set up

```{r, include = F}
knitr::opts_chunk$set(echo=T, warning=FALSE, message=FALSE,
                      fig.width=12, fig.height=8)
```


```{r}
rm(list=ls())
library(magrittr)
library(stringr)
library(knitr)
library(tidyverse)
library(RMySQL)
library(directlabels)
theme_set(ggthemes::theme_few())
```

## Read data

```{r}
all_qs <- read_csv("../data/childes_questions.csv")
```

## Descriptives

```{r}
all_qs %>% distinct(corpus) %>% kable()
```

```{r}
all_qs %>% 
  distinct(child) %>% 
  nrow() %>% 
  paste("n children =", .)
```

```{r}
all_qs %>% 
  ungroup() %>% 
  group_by(speaker_clean, gloss_clean, length_transcript) %>% 
  summarise(count = n()) %>% 
  ggplot(., aes(x = gloss_clean, y = count)) +
  geom_bar(stat = "identity") +
  labs(x = "Question Word") +
  facet_wrap(~speaker_clean)
```

```{r}
all_qs %>% 
  group_by(speaker_clean, parent_type) %>% 
  summarise(count = n())
```

```{r}
all_qs %>% 
  filter(speaker_clean == "child") %>% 
  select(gloss_clean, sentgloss, age_months) %>% 
  group_by(gloss_clean) %>% 
  sample_n(., size = 2) %>% 
  knitr::kable()
```

## Exploratory analyses

Count different kinds of questions for each kid and plot as a function of age.

```{r}
all_qs %>% 
  filter(speaker == "CHI") %>% 
  group_by(child_clean, age_months, gloss_clean, length_transcript) %>% 
  summarise(freq = n()) %>% 
  filter(freq > 0) %>% 
  mutate(prop = freq / length_transcript) %>%
  ggplot(., aes(x = age_months, y = prop, color = gloss_clean)) +
  geom_smooth(method = "loess", se = F) +
  labs(color = "question word", y = "Prop. of Utterances") +
  langcog::scale_color_solarized() +
  facet_wrap(~child_clean, ncol = 7) +
  theme(legend.position = "bottom") +
  ggtitle("Children's questions")
```

Count different kinds of questions for each parent and summarise over development

```{r}
all_qs %>% 
  filter(speaker_clean == "parent") %>% 
  group_by(child_clean, age_months, gloss_clean, length_transcript) %>% 
  summarise(freq = n()) %>% 
  filter(freq > 0) %>% 
  mutate(prop = freq / length_transcript) %>% 
  ggplot(., aes(x = age_months, y = prop, color = gloss_clean)) +
  geom_smooth(method = "loess", se = F) +
  labs(color = "question word") +
  langcog::scale_color_solarized() +
  facet_wrap(~child_clean, ncol = 7) +
  theme(legend.position = "bottom") +
  ggtitle("Parents' questions")
```

Now get the average frequency for each question type at each month of development.

```{r}
all_qs %>% 
  filter(gloss_clean %in% c('what', 'where', 'why')) %>% 
  group_by(child_clean, age_months, gloss_clean, speaker_clean, length_transcript) %>% 
  summarise(freq = n()) %>% 
  filter(freq > 0) %>% 
  mutate(prop = freq / length_transcript) %>% 
  group_by(age_months, gloss_clean, speaker_clean) %>% 
  summarise(m = mean(prop, na.rm = T),
            n = n()) %>% 
  ggplot(., aes(x = age_months, y = m, color = gloss_clean)) +
  geom_point(aes(size = n), alpha = 0.3) +
  geom_smooth(se=F, method = "loess", size = 2) +
  xlim(10, 68) +
  labs(y = "Prop. of Utterances", x= "Age (months)") +
  guides(color = F) +
  langcog::scale_color_solarized() +
  geom_dl(aes(label = gloss_clean), method=list("last.points")) +
  facet_wrap(~speaker_clean, scales = "free") +
  theme(legend.position = "top")
```

### Gender analysis

```{r}
gender_ms <- all_qs %>% 
  group_by(child_clean, gloss_clean, speaker, gender) %>% 
  summarise(freq = n()) %>%
  filter(freq > 0, gender %in% c("male", "female")) %>% 
  group_by(gloss_clean, speaker, gender) %>% 
  langcog::multi_boot_standard(column = "freq", empirical_function = "mean") 
```

```{r}
gender_ms %>% 
  ggplot(., aes(x = gloss_clean, y = mean, color = gender)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), 
                  position = position_dodge(width = 0.3), size = 0.8) +
  labs(y = "Avg. Freq", x= "Question Type") +
  langcog::scale_color_solarized() +
  facet_wrap(~speaker, scales = "free") +
  theme(legend.position = "top")
```

No gender differences in question words produced by kid or by mom.

### Context analyses

Here we will attempt to operationalize cost and reward of a behavior to get the value for a question.

Spread participants variable, so we can code for the number of people in the conversation.

```{r}

```

### Complexity of utterance that each question word occurs in

Use some string functions to get the word count for each utterance.

```{r}
all_qs %<>% 
  rowwise() %>% 
  mutate(sentence_len = length(strsplit(sentgloss,' ')[[1]]))
```

Now summarise the mean sentence length for each question type for each kid and speaker

```{r}
sentence_len_ms <- all_qs %>% 
  ungroup() %>% 
  group_by(age_months, speaker_clean, child_clean, gloss_clean) %>% 
  summarise(m.ss = median(sentence_len), 
            n = n()) %>% 
  group_by(age_months, speaker_clean, gloss_clean) %>% 
  summarise(m = mean(m.ss), 
            n = n()) 
```

```{r}
sentence_len_ms %>% 
  filter(gloss_clean %in% c("where", "what", "why")) %>% 
  ggplot(data = ., aes(x = age_months, y = m, color = gloss_clean)) +
  geom_point(aes(size = n), alpha = 0.3) +
  geom_smooth(se=F, method = "loess", size = 2) +
  xlim(10, 68) +
  labs(y = "Avg. Sentence Length", x = "Age (months)") +
  guides(color = F) +
  langcog::scale_color_solarized() +
  geom_dl(aes(label = gloss_clean), method=list("last.points")) +
  facet_wrap(~speaker_clean) +
  theme(legend.position = "top")
```

### Single question word utterances

```{r}
single_words <- all_qs %>% 
  filter(sentence_len == 1) %>% 
  group_by(speaker_clean, child_clean, gloss_clean, age_months, length_transcript) %>% 
  summarise(freq = n()) %>% 
  mutate(prop = freq / length_transcript) %>% 
  group_by(speaker_clean, gloss_clean, age_months) %>% 
  summarise(m.ss = median(prop), 
            n = n())
```

```{r}
single_words %>% 
  filter(gloss_clean %in% c("where", "what", "why")) %>% 
  ggplot(data = ., aes(x = age_months, y = m.ss, color = gloss_clean)) +
  geom_point(aes(size = n), alpha = 0.3) +
  geom_smooth(se=F, method = "loess", size = 2) +
  xlim(10, 68) +
  guides(size = F) +
  labs(y = "Prop. of Utterances", x = "Age (months)") +
  langcog::scale_color_solarized() +
  facet_wrap(~speaker_clean) +
  theme(legend.position = "top")
```

### Fathers compared to mothers on wh-questions

```{r}

```


### Other analyeses to do:

- 
